GPU Acceleration of Edge-Based Motion Detection and Machine Learning-Aided Facial Recognition with NVIDIA CUDA

### 2015-12-16 documents/paper/paper.tex
Corrected typos in paper

\section{Introduction}
Graphics processing units (GPUs) are rapidly gaining popularity as a platform for parallelized computations on massive sets of data. Since much of the computations in image processing and computer vision are easily parallelized, graphics operations on GPUs achieve significant speedups compared to those done on their serial, CPU counterparts. Further, SDKs like the NVIDIA CUDA framework provide developers easy APIs to take advantage of the parallel computing power of GPUs. We take full advantage of the computational benefits of GPUs by implementing edge detection and motion detection algorithms in CUDA C, and making use of existing CUDA libraries for our facial recognition algorithm.
\par In this paper, we first detail the theory for our edge detection, motion detection, and facial recognition algorithms in Sections \ref{edge-detection-section}, \ref{motion-detection-section}, and \ref{facial-recognition-section}, respectively. At the end of Sections \ref{edge-detection-section} and \ref{motion-detection-section}, we describe our GPU code implementation of these algorithms with NVIDIA CUDA. At the end of Section \ref{facial-recognition-section}, we comment on the performance we achieve with a prebuilt, CUDA-based OpenCV GPU computation library, as opposed to that we achieve with a custom CUDA implementation as in Sections \ref{edge-detection-section} and \ref{motion-detection-section}. We present speedup results achieved with our CUDA implementation with respect to a reference serial, CPU implmentation in Section \ref{results}. Finally, we conclude in Section \ref{conclusion}.

@@

\subsection{Non-Maximum Suppression and Selective Thresholding}
Given the gradient magnitude and direction of the Gaussian-filtered image, the final step of the edge detection algorithm is to determine which pixels should be selected as edge pixels, and which should be rejected. We define two procedures, \textit{non-maximum suppression} and \textit{selective thresholding}, to gain a reasonably accurate edge map from the gradient.
\par Non-maximum suppression selects and rejects edge pixels according to the following criteria:
\begin{enumerate}
	\item Pixels whose gradient magnitude is below a user-defined low threshold, $t_l$, is immediately rejected
	\item If the gradient magnitude of a pixel is greater than that of the pixels in either direction of the gradient angle of that pixel, then the pixel is accepted if its gradient magnitude is greater than a user-defined high threshold, $t_h$
\end{enumerate}
More precisely, a pixel at $(i, j)$ with gradient magnitude $G_{i, j} > t_l$ and angle $\theta$ is accepted if any of the following hold true:
\begin{itemize}
	\item $(-\frac{\pi}{6} < \theta < \frac{\pi}{6} \vee -\pi < \theta < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \theta < \pi) \wedge G_{i, j} > G_{i, j \pm 1}$
	\item $(\frac{\pi}{6} < \theta < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \theta < -\frac{2 \pi}{3}) \wedge G_{i, j} > G_{i \pm 1, \pm j}$
	\item $(\frac{\pi}{3} < \theta < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \theta < -\frac{\pi}{3}) \wedge G_{i, j} > G_{i \pm 1, j}$
	\item $(\frac{2 \pi}{3} < \theta < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \theta < -\frac{\pi}{6}) \wedge G_{i, j} > G_{i \pm 1, \mp j}$
\end{itemize}
The selective thresholding technique extends on the procedure of non-maximum suppression. For any pixel that passes crtiteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel. If any pixel within the box exceeds the low threshold $t_l$, then the pixel is accepted. This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satify the criterion themselves, in order to create more continuous edge lines.
\begin{figure}[]
	\centering
	\includegraphics[width=0.45\textwidth]{pcb_edges.jpg}
	\caption{Output of edge detection algorithm after non-maximum suppression and selective thresholding on the image of Figure \ref{pcb-original}, with parameters $t_l = 70$ and $t_h = 80$}
    \label{pcb-edges}
\end{figure}
\begin{algorithm}[h]
	\smaller
	\label{non-maximum-suppression-selective-thresholding-algorithm}
	\caption{Non-maximum suppression and selective thresholding: pseudocode for a serial, CPU implementation}

@@

\subsection{Facial Recognition}
The below results reflect the speedup achieved using OpenCV's object detection framework.
\begin{table}[H]
	\small
	\centering
	\caption{Speedup of Facial Recongition}
	\label{facial-recognition-speedup}
	\begin{tabular}{llllll}
	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
	 $100 \times 100$ & $0.067883$ & $0.061132$ & $1.110433$ \\
	 $500 \times 500$ & $0.226192$ & $0.094812$ & $2.385690$ \\
	 $1000 \times 1000$ & $0.811302$ & $0.199005$ & $4.076792$ \\
	 $2000 \times 2000$ & $2.849201$ & $0.562820$ & $5.062366$ \\
	 $3000 \times 3000$ & $6.236422$ & $1.218879$ & $5.116523$ \\
	 $4000 \times 4000$ & $10.794892$ & $2.142449$ & $5.038576$ \\
	\end{tabular}
\end{table}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{facial_recognition_speedup_graph.jpg}
	\caption{Visualization of facial recognition speedup versus pixel count}
    \label{facial-recognition-speedup-graph}
\end{figure}

>>>>

\section{Introduction}
Graphics processing units (GPUs) are rapidly gaining popularity as a platform for parallelized computations on massive sets of data. Since much of the computations in image processing and computer vision are easily parallelized, graphics operations on GPUs achieve significant speedups compared to those done on their serial, CPU counterparts. Further, SDKs like the NVIDIA CUDA framework provide developers easy APIs to take advantage of the parallel computing power of GPUs. We take full advantage of the computational benefits of GPUs by implementing edge detection and motion detection algorithms in CUDA C, and making use of existing CUDA libraries for our facial recognition algorithm.
\par In this paper, we first detail the theory for our edge detection, motion detection, and facial recognition algorithms in Sections \ref{edge-detection-section}, \ref{motion-detection-section}, and \ref{facial-recognition-section}, respectively. At the end of Sections \ref{edge-detection-section} and \ref{motion-detection-section}, we describe our GPU code implementation of these algorithms with NVIDIA CUDA. At the end of Section \ref{facial-recognition-section}, we comment on the performance we achieve with a pre-built, CUDA-based OpenCV GPU computation library, as opposed to that we achieve with a custom CUDA implementation as in Sections \ref{edge-detection-section} and \ref{motion-detection-section}. We present speedup results achieved with our CUDA implementation with respect to a reference serial, CPU implementation in Section \ref{results}. Finally, we conclude in Section \ref{conclusion}.

@@

\subsection{Non-Maximum Suppression and Selective Thresholding}
Given the gradient magnitude and direction of the Gaussian-filtered image, the final step of the edge detection algorithm is to determine which pixels should be selected as edge pixels, and which should be rejected. We define two procedures, \textit{non-maximum suppression} and \textit{selective thresholding}, to gain a reasonably accurate edge map from the gradient.
\par Non-maximum suppression selects and rejects edge pixels according to the following criteria:
\begin{enumerate}
	\item Pixels whose gradient magnitude is below a user-defined low threshold, $t_l$, is immediately rejected
	\item If the gradient magnitude of a pixel is greater than that of the pixels in either direction of the gradient angle of that pixel, then the pixel is accepted if its gradient magnitude is greater than a user-defined high threshold, $t_h$
\end{enumerate}
More precisely, a pixel at $(i, j)$ with gradient magnitude $G_{i, j} > t_l$ and angle $\theta$ is accepted if any of the following hold true:
\begin{itemize}
	\item $(-\frac{\pi}{6} < \theta < \frac{\pi}{6} \vee -\pi < \theta < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \theta < \pi) \wedge G_{i, j} > G_{i, j \pm 1}$
	\item $(\frac{\pi}{6} < \theta < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \theta < -\frac{2 \pi}{3}) \wedge G_{i, j} > G_{i \pm 1, \pm j}$
	\item $(\frac{\pi}{3} < \theta < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \theta < -\frac{\pi}{3}) \wedge G_{i, j} > G_{i \pm 1, j}$
	\item $(\frac{2 \pi}{3} < \theta < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \theta < -\frac{\pi}{6}) \wedge G_{i, j} > G_{i \pm 1, \mp j}$
\end{itemize}
The selective thresholding technique extends on the procedure of non-maximum suppression. For any pixel that passes criteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel. If any pixel within the box exceeds the low threshold $t_l$, then the pixel is accepted. This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satisfy the criterion themselves, in order to create more continuous edge lines.
\begin{figure}[]
	\centering
	\includegraphics[width=0.45\textwidth]{pcb_edges.jpg}
	\caption{Output of edge detection algorithm after non-maximum suppression and selective thresholding on the image of Figure \ref{pcb-original}, with parameters $t_l = 70$ and $t_h = 80$}
    \label{pcb-edges}
\end{figure}
\begin{algorithm}[h]
	\smaller
	\label{non-maximum-suppression-selective-thresholding-algorithm}
	\caption{Non-maximum suppression and selective thresholding: pseudocode for a serial, CPU implementation}

@@

\subsection{Facial Recognition}
The below results reflect the speedup achieved using OpenCV's object detection framework.
\begin{table}[H]
	\small
	\centering
	\caption{Speedup of Facial Recognition}
	\label{facial-recognition-speedup}
	\begin{tabular}{llllll}
	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
	 $100 \times 100$ & $0.067883$ & $0.061132$ & $1.110433$ \\
	 $500 \times 500$ & $0.226192$ & $0.094812$ & $2.385690$ \\
	 $1000 \times 1000$ & $0.811302$ & $0.199005$ & $4.076792$ \\
	 $2000 \times 2000$ & $2.849201$ & $0.562820$ & $5.062366$ \\
	 $3000 \times 3000$ & $6.236422$ & $1.218879$ & $5.116523$ \\
	 $4000 \times 4000$ & $10.794892$ & $2.142449$ & $5.038576$ \\
	\end{tabular}
\end{table}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{facial_recognition_speedup_graph.jpg}
	\caption{Visualization of facial recognition speedup versus pixel count}
    \label{facial-recognition-speedup-graph}
\end{figure}
